\chapter{Tools}
\label{section:tools}


\section{Loop macros}

\subsection{d-dimensional loops}

I wrote Peano to support arbitrary spatial dimensions.
By default, I build only 2d and 3d versions of the code/core, but in principle
you can write also simulations over 4d or 5d domains, e.g.
The choice of a particular dimension d is done at compile time via the
flag \texttt{-DDimension=d}.
As everything is generic w.r.t.~dimensions, I have many places in the code where
I need d-dimensional loops.


Peano's utility directory/namespace \texttt{peano4::utils} offers a header
\texttt{Loop} which comprises a set of macros that support d-dimensional loops.
Details can be found in the source code documentatation within the file or
online, but the collection of macros allows you to write code alike

\begin{code}
  dfor(k,20) { // d-dimensional loop over 20x20x20x... patch
    // There is now a d-dimensional vector k over integers which holds the 
    // loop index. So we can write for example
    const tarch::la::Vector<Dimensions,double> somePosition = 
      offset + k.convertScalar<double>() * h;
    [...]
  }
\end{code}

\noindent
There are also macros for loops skipping one dimension, or macros for particular
iteration ranges (2,3,4) which are highly optimised through lookup tables.
Finally, \texttt{Loop} hosts z-loops, i.e.~loops which meander forth and back
through an array similar to the Peano curve.


To use the d-dimensional loops, you have to include
\begin{code}
#include "peano4/utils/Loop.h" 
\end{code}


For Cartesian meshes, arrays, the file also offers a routine
\texttt{peano4::utils::dLinearised} which converts the d-dimensional loop
counter into one integer.
Again, I do offer variants of this routine which are optimised to particular
choices of d or array ranges.

\subsection{d-dimensional exclusive loops}

There's a variation of the d-dimensional loops which loops over all dimensions
except one.
I name this loop type \texttt{dfore}.
Its usage is straightforward: 
You again specify a counter name and you pass in the loop range. 
Let \texttt{k} be the counter name and \texttt{N} the range.
You furthermore specify a dimension and you fix the value for this dimension.

The loop will run over a \texttt{N} $\times $ \texttt{N} $\times $ 1 $\times
$\texttt{N} range with the 1 being in the specified direction.
The loop counter is a vector over integer named \texttt{k}.
Its entry along the 1-dimension is fixed to the default value specified in the
macro.


\section{Synchronisation}

\subsection{Shared memory synchronisation}

In the tradition of Intel's original TBB version, I offer an abstraction over
critical sections aka exclusive code blocks.
That is blocks that can only be executed by one thread at a time.

To use it, you define a \texttt{BooleanSemaphore} object.
It should be global (or static) and is an abstraction around C++/TBB/OpenMP.
To lock it, you can do so manually, but actually I offer a \texttt{Lock} object
which locks the semaphore and automatically frees it once the object is
destroyed.

\subsection{Distributed memory synchronisation}

I also offer a boolean semaphore accross all MPI ranks. This one is quite
expensive, as it has to lock both all threads and all ranks and thus send
quite some messages forth and back. 

The usage is similar to the shared memory lock, but you have to give the
semaphore a unique name (you might construct a string using the
\texttt{\_\_LINE\_\_} and \texttt{\_\_FILE\_\_} macro). 
It is the name that \Peano\ uses to ensure that the semaphore is unique accross
all ranks.
In this way, the semaphore is similar to an OpenMP named critical section.
