\chapter{Introducing a new pre-defined action set}
\label{section:new-action-set}

This chapter discusses how you make an action set (set of event implementations)
of your code available to other projects as a pre-defined action set.
As a result of such a refactoring, others can inject your ideas
straightforwardly into their project.
The name refactoring suggests that I highly recommend that you start from a
working action set that you've written manually and extensively tested.

\section{Prepare classes}

Create a new subclass of \texttt{peano4.solversteps.ActionSet}

\begin{code}
import peano4.solversteps.ActionSet

class MyActionSet(peano4.solversteps.ActionSet):
  def __init__(self):
    pass
\end{code}

\noindent
and ensure that Python finds the class in its path.


\section{Implement new action set}

To implement the new action set, you have to tell Peano what the function bodies
look like.
Before we do so, we instruct the toolkit whether this is an action set that
users should later on edit; 
that is, your new predefined action set serves as blueprint.
Alternatively, your action set can be read-only and users shall not modify it.
In this case, Peano will place it in the observers' directory and always
regenerate it automatically.


The first step hence is to give your action set a clear name (which will also be
used as class name part for the generated file) and to provide this information:


\begin{code}
  def get_action_set_name(self):
    // My default implementation that I use in the toolboxes is this following
    // line. You are however allowed to use your own, bespoke name
    // return __name__.replace(".py", "").replace(".", "_")
    return "MyName"

  def user_should_modify_template(self):
    return False
\end{code}


\noindent
The actual events are generated through the operation bodies.
All these bodies are queries through one operation. 
The query's arguments clarifies which operation body is requested

\begin{code}
  def get_body_of_operation(self,operation_name):
    # An examples where we only do something for :
    if operation_name == ActionSet.OPERATION_CREATE_CELL:
      return "mycell.add(...)\n"
    return "// Nothing to implement\n"
\end{code}

\noindent
and its values are those constants in \texttt{Mapping} which start with
\texttt{OPERATION\_}.


There are further routines to control the refinement et al.
These routines are self explaining if you study the base class/interface.


More interesting is the question how to make the actual method bodies use grid
information.
For this, the canonical way is that you extend the constructor such that an
object of your action set always ``knows'' which data they work on. 
The interfaces of all action sets are always the same, so it should be
straightforward to make your generated method bodies use the grid data once the
constructor is told what they are.


\section{Use your new action set}

To use the action set, you create your algorithmic steps.
Once created, you also create and instance of your action set and you add it to the
step with \texttt{add\_action\_set}.
The order in which action sets are added is important.
See \texttt{add\_action\_set}'s documentation of information on this.



\section{Conventions}

By default, the Peano toolkit equips each action set with a couple of
features/attributes:

\begin{itemize}
  \item Every action set has a static attribute \texttt{\_log} which is an instance
  of \texttt{tarch::logging::Log}. That is, you can use Peano's logging commands
  straightaway.
\end{itemize}

